---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file: `devtools::build_readme()` is handy for this. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gatewayr

<!-- badges: start -->
<!-- badges: end -->

The goal of gatewayr is to interface with the GATEWAy database and run simple analyses.

## Installation

You can install the development version of gatewayr like so:

``` r
devtools::install_github("emilio-berti/gatewayr")
```

## Compute network and community metrics for a study site

Here, we will calculate common network metrics and the jaccard similarity of species between foodwebs of the same study site.
As example, we will use the `Adirondack lakes` as study site.
First, we subset the GATEWAy database to keep only the food webs within this study site.
This is achieved using `subset(name, pattern)`, where `name` is the name of a column in GATEWAy (in this case `"study.site"`) and pattern is `"Adirondack lakes"`.
There is no need to load the GATEWAy database, as this is comes with teh gatewayr package itself.

```{r subset}
library(gatewayr)

d <- subset("study.site", "Adirondack lakes")
```

We first calculate the Jaccard similarity of species between all pairs of food webs using `species_jaccard()`:

```{r, jaccard}
jaccard <- species_jaccard(d)
jaccard[1:3, 1:3]
image(jaccard, col = hcl.colors(100, "DarkMint"))
```

To calculate the network metrics for a food web we can use `network_metrics()`:
```{r metrics}
A <- adjacency(d[d$foodweb.name == d$foodweb.name[1], ])
net <- network_metrics(A)
net
```

The function `network_dissimilarity()` calls this functions for each food web in the dataset and calculate the euclidean distance between foodwebs in PCA space:

```{r dissimlarity}
network_distance <- network_dissimilarity(d)
image(network_distance, col = hcl.colors(100, "Zissou 1"))
```

## Analysis of dissimilarities

Here is a simple analysis of the relationship between species similariy and network dissimilarity.

```{r plot}
fit <- loess(as.vector(network_distance) ~ as.vector(jaccard))
lm <- lm(as.vector(network_distance) ~ as.vector(jaccard))

plot(
  jaccard, network_distance,
  pch = 20,
  cex = .6,
  col = adjustcolor("dodgerblue3"),
  frame = FALSE,
  xlab = "Species similarity (Jaccard)",
  ylab = "Network dissimilarity (PCA distance)"
)
abline(lm, lw = 3, col = "grey20")

summary(
    lm(
    scale(as.vector(network_distance)) ~ 
    -1 + scale(as.vector(jaccard))
  )
)
```

## Bootstrapping

The models above have many points, however we only have 50 food webs, which is our true sample size.
In other words, we have pseudo-replication.
Bootstrapping is a crude way to solve this problem.

```{r bootstrap}
plot(
  c(0, 0.55), c(0, 11),
  cex = 0,
  frame = FALSE,
  xlab = "Species similarity (Jaccard)",
  ylab = "Network dissimilarity (PCA distance)"
)

n <- 1000  # bootstrap samples
slopes <- rep(NA, n)
p_values <- rep(NA, n)
for (i in seq_len(n)) {
  # first set of food webs
  first_set <- sample(
    rownames(jaccard),
    floor(nrow(jaccard) / 2)
  )
  # second set of food webs
  second_set <- sample(
    setdiff(rownames(jaccard), first_set),
    floor(nrow(jaccard) / 2)
  )
  comparisons <- cbind(first_set, second_set)
  # subset the whole dataframe to keep only 25 pairwise comparisons
  boot <- data.frame()
  for (j in seq_len(nrow(comparisons))) {
    boot <- rbind(
      boot,
      data.frame(
        jaccard = jaccard[comparisons[j, 1], comparisons[j, 2]],
        network = network_distance[comparisons[j, 1], comparisons[j, 2]]
      )
    )
  }

  with(boot, points(
    jaccard, network,
    pch = 20,
    cex = .6,
    col = adjustcolor("dodgerblue3", alpha.f = .3)
  ))

  lm <- with(boot, lm(network ~ jaccard))
  abline(lm, lw = 2, col = adjustcolor("grey20", alpha.f = .2))
  lm <- with(boot, lm(scale(network) ~ -1 + scale(jaccard)))
  slopes[i] <- coef(lm)
  p_values[i] <- anova(lm)$`Pr(>F)`[1]
}

summary(slopes)
hist(slopes, breaks = 10, main = "", xlab = "Slopes")
```


