---
title: "api-usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{api-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gatewayr)
```

# Access Food Web metadata
The `gateway_foodwebs()` accesses the food web metadata:
  
  - Name
  - Ecosystem type
  - GPS coordinates
  - Altitude & depth
  - Sampling time

```{r foodwebs}
gateway_foodwebs()
```
The optional arguments _ecosystemType_ and _xmin_, _xmax_, _ymin_, _ymax_ can be used to filter the database to retain only the food webs from one ecosystem type or within a bounding box.
```{r foodwebs-filtering}
gateway_foodwebs(ecosystemType = "lakes")
gateway_foodwebs(ecosystemType = "lakes", xmin = 5)
```

# Access Taxon List
The `gateway_taxa()` accesses the taxon information:
```{r taxa}
gateway_taxa()
```
The optional arguments _taxonRank_ and _taxonomicStatus_ can be used to retain only taxa of a specific rank (e.g., _species_ or _family_) or taxonomic status (e.g., _accepted_ or _synonym_):
```{ taxa-filtering}
gateway_taxa(taxonRank = "species")
gateway_taxa(taxonRank = "species", taxonomicStatus = "accepted")
```
